---
title: "learningR"
author: "Yetunde"
date: "11/25/2021"
output: html_document
---

```{r}
34 + 5
```

```{r}
45 * 2
```

```{r}
names <- c("John", "Harrison", "Bolu")
names
```

```{r}
x<- 2+3
y<-(5*11)/4-7
s<-7^3
x
y
s
```
sequence and matrix
```{r}
x<- 1:50
x
length(x)
X<-seq(1,9,by =2)
X

name<- matrix(1:16, nrow = 4, ncol = 4)
name


dim(name)

numb<- c(1:10)
dim(numb) <- c(2,5)
numb
```

```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
colnames(m) <- c("a", "b")
rownames(m) <- c("c", "d")
m
```

```{r}
x <- 1:4
y <- 5:8

M <- cbind(x, y)

Q <- rbind(x,y)

colnames(M) <- c("col1", "col2")
rownames(M) <- c("row1", "row2", "row3", "row4")
M

colnames(Q) <- c("col1", "col2", "col3", "col4")
rownames(Q) <- c("row1", "row2")
Q

```
list
```{r}
X <- list(a =1, b = "c", c =3)
X
X$b
```
factor
```{r}
gender <- c("male", "male", "female", "female", "male", "female", "male")
as.factor(gender)
sum(is.na(gender))
is.nan(gender)
```

```{r}
mis <- c(1,3,4,NA,5.7)
mean(is.na(mis))
```
data frame
```{r}
emp.data <- data.frame(emp_id = c(1:5), 
emp_name = c("Rick", "Dan", "Michelle", "Ryan", "Gary"),
salary = c(623.3, 515.2, 611.0, 729.0, 843.25))

emp.data[3,]
emp.data$emp_name[2]
```

```{r}
x <- 5:7
names(x)

names(x) <- c("five", "six", "seven")
x
```
```{r}
x <- c("a", "b", "c", "d")
x[2]
x[1:4]
x[x>"b"]
```

```{r}
x <- matrix(1:6, 2,3)
x
x[1,2]
x[2,1]
x[1, 2, drop= FALSE]
x[1,, drop= FALSE]
```
Subsetting a matrix
```{r}
x <- matrix(1:12, 3,4)
x
x[2:3, 2:4]
```

```{r}
Y <- matrix(1:16, 4,4)
colnames(Y) <- c("col1", "col2", "col3", "col4")
Y
Y[1,2]
Y[1:3, 3:4]
```
Data frame

```{r}
file <- read.csv("~/Documents/R/Fertility (1).csv", sep=",", header=TRUE, col.names = c(1,2,3,4,5,6,7,8,9))
file
sum(is.na(file))
mean(is.na(file))
na.omit(file)
```

```{r}
file <- read.csv("~/Documents/R/who_suicide_statistics_semi.csv", sep = ";", strip.white = TRUE)
file
sum(is.na(file))
median(is.na(file))
```


```{r}
auto <- read.table("https://s3.amazonaws.com/assets.datacamp.com/blog_assets/test.txt", header= FALSE)
head(auto[,2:3])
```

```{r}
hs <- read.csv("~/Documents/R/house.csv", sep=",")
hs
boxplot(Price~Status, data = hs, main = "price")

plot(hs$SQFT, hs$Price, xlab="SQFT", ylab = "Price", main = "ScatterPlot of Price and SQFT")

hist(hs$Price, xlab="Price", main="Histogram of Price")
```

```{r}
cancer <- read.csv("~/Downloads/cancer.csv", sep = ",")
cancer

boxplot(radius_mean~diagnosis, data = cancer, main = "mean")
hist(cancer$radius_mean, xlab = "dignosis_mean", main = "Histogram of radius mean")
```

```{r}
library(readdxl)
library(tidyverse)
```

loading mbta dataset, skipping the first row from excel sheet one pick col 2 to 60
```{r}
mbta <- read_excel("~/Downloads/mbta.xlsx", sheet = 1, skip = 1, range = cell_cols(2:60))
mbta
```

gathering data from january 2007, t0 2011, january
```{r}
mbta <- mbta %>% gather('2007-01':'2011-10', key = "year", value = "passenger")
mbta

seperating the year into two column
```{r}
mbta <- mbta %>% separate(year, into = c("year", "month"))
mbta
```

Transforming rows ro column by using pivot_wider function 
```{r}
mbta <- mbta %>% pivot_wider(names_from = mode, values_from = passenger)
mbta
```

```{r}
mbta <- mbta %>% .[, c(1:2, 6:8)]
mbta
                   
```

```{r}
mbta <- mbta %>% gather('Commuter Rail':'Light Rail', key = "rail_type", value = passengers)
mbta
```